Bridging Frontend and Backend: Implementing Sigma Protocol in WebAssembly and Rust from a Single Coq Formalisation

### 2024-06-10 paper/IEEES/paper.tex
Tmp commit

\usepackage{listings}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{tikz}
\usetikzlibrary{positioning, arrows.meta}
\usepackage{url}
\usepackage{xcolor}
\usepackage{soul}
\usepackage{lstcoq}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}
\title{Hope}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
  \st{Despite the recent growth of zero-knowledge succinct argument 
  of knowledge (ZkSNARK)}, 
  $\Sigma$-protocol remains a popular choice in electronic voting to maintain privacy 
  and verifiability of elections and still used in prominent e-voting systems such as 
  Helios, Belenois, SwissPost, etc. However, 
  the existing implementations of sigma protocol have two 
  shortcomings: (i) the same code is written twice, once for the front-end (JavaScript/TypeScript)
  and once for the back-end (Python/OCaml/Java), which is time-consuming but more importantly error-prone, 
  and (ii) none of these implementations offer any security guarantees.

@@

\subsection{Outline}
Outline goes here.

\section{Sigma Protocol}\label{sigma_protocol}
Despite the recent growth and popularity of zero-knowledge succinct 
argument of knowledge (ZKSNARK), $\Sigma$-protocol remains a
leading cryptographic proof system in privacy-preserving voting systems. 
$\Sigma$-protocol is used in Helios --a widely popular voting system-- \cite{adida2008helios}, 
Belenois --used in French elections for overseas 
voters\footnote{VVFE, used in French elections, is a derivative of (part of) Belenios and 
inherited all the zero-knowledge proofs from Belenois.}-- \cite{cortier2023french}, and 
SwissPost --used in Swiss elections for overseas voters-- voting 
systems \cite{10.1007/978-3-031-15911-4_4} (\textcolor{red}{Look for more examples}). 
Moreover, recently zero-knowledge proof community started a movement to standardise
the sigma protocol \cite{ZKProof}.
One of the main reasons for its popularity in privacy-preserving voting systems is its
well-understood security.  Other application of sigma protocol 
include identification schemes, (\textcolor{red}{cite more work}).

%\textbf{Introduce here that sigma protocol is efficient and simple and all other details about sigma protocol.} 
Zero-knowledge proofs were first studied by Goldwasser, Micali, and Rackoff \cite{10.1145/22145.22178} and 
are possible for all problems in $NP$ \cite{10.1145/116825.116852}. Recall that $NP$ is the class of 
problems that have efficient verifiers; that is, for a given problem in the $NP$ class, there is 
a polynomial-time algorithm that can verify whether a given solution, or witness, to the problem is correct or not.
More formally, for a given $NP$-relation $R$ and a statement $x$, zero-knowledge proof allows a prover 
$P$ to convince a verifier $V$ that they posses a witness $w$ --polynomial in length of $x$-- such 
that ($x$, $w$) $\in$ $R$. To achieve this, $P$ and $V$ engage in a 
interactive protocol and  at the end of the protocol $V$ either accepts or rejects the proofs.
\textit{Sigma protocol}, shown in the figure \ref{fig:sig_picture}, is an efficient three move zero-knowledge proof where the 
verifier is assumed to be honest and characterised by completeness, (special) soundness, and 
(special honest-verifier) zero-knowledge. More formally: 
\begin{itemize}
  \item \textbf{Completeness:}
  when $P$ and $V$ follow the protocol, $V$ always accepts. 
  In other words, if $P$ knows $w$ for $x$ such that ($x$, $w$) $\in$ $R$,  
  then $P$ can successfully construct a proof $(a, c, r)$ that always passes the verification equation, 
  i.e., $V$ accepts.

\item \textbf{Special soundness:}
there exists a polynomial-time extractor $E$ which given 
two accepting transcripts $(a, c, r)$ and $(a, c', r')$ for $x$, 
it can extract a witness $w$ such that $(x, w) \in R$. In other words, 
if $P$ does not know $w$ for $x$ then it cannot construct a proof 
(a; c; r) that will passes the verification check. 
\item \textbf{Special honest-verifier zero-knowledge proof:}
There exists a polynomial-time simulator $S$ that, for any 
$v \in L_R$ where $L_R = \{x  \mid (x, w) \in R\}$,
generates an accepting transcript $(a, c, r)$ 
without using the witness $w$. 
This transcript is indistinguishable from one produced by an 
honest conversation between $P$ and $V$ using $w$, 
satisfying $(v,w) \in R$.

@@

Sigma protocol was first defined by Ronald Cramer \cite{cramer1996modular} 
and the first efficient sigma protocol was introduced by Schnorr in \cite{schnorr1991efficient}. 
Eventhough there exists many sigma protocols, the Schnorr protocol is used extensively in privacy-preserving voting systems, e.g, Helios, Belenois, SwissPost, etc. Therefore, 
our formalisation is geared towards the Schnorr protocol. 
Moreover, we also formalise the composition of (i) $n$ \textit{Parallel} Schnorr, 
(i) $n$ \textit{And} Schnorr, (iii) $n$ \textit{Or} Schnorr, and (iv) $n$ \textit{Eq} Schnorr protocol. 
Ideally, given the abstraction of our formlisation (\textit{vector space}), we could have 
abstracted the sigma protocol and their composition in a generic way \cite{10.1007/978-3-642-02384-2_17},
but we decided to focus on the Schnorr protocol and its composition to ensure familiar APIs 
for e-voting in WebAssembly and Rust. We also formalise Okamoto protocol as a subprotocol 
for Neq composition (\textbf{Berry: does it sound okay?}). However, we have not seen 
much use of Neq in the privacy-preserving voting systems, at least in the existing voting systems.

\section{The Coq Theorem Prover}
\textcolor{red}{FIXME:this seems too much of tutorial Coq so let me know your thoughts. We 
can explain MetaCoq here as well and the extraction to Rust and WebAssembly
so that it justifies this tutorial}
The Coq theorem prover \cite{the_coq_development_team} is a computer program that interacts with users, 
enabling them to encode mathematical definitions, express specifications (true statements) about 
their definitions, and formally prove that the definitions imply the specifications. 
A user first defines their mathematical object (definition) and then figure 
out some  true statements (specifications) about 
that object. Once the user has definition 
and specifications, they need to prove 
that the definition satisfies the specifications 
(proof writing). In general, amongst these three steps, 
the proof writing step is the most challenging and important step. 
During the proof writing, an user interacts with the Coq theorem prover 
via \emph{tactics}, a domain specific language 
to ease the proof writing. Even though the Coq theorem 
provides some amount of automation and proof 
search, most of the time it requires human assistance to finish the proof. 
It is one the most popular and robust theorem, and
one of reasons for its popularity is 
dependent types, that makes it possible to encode
any mathematical statement in its logic.
It has been used to verify many real world software projects and mathematical 
artefacts. For example, Coq has been used to 
formally verify CompCert \cite{10.1145/1111037.1111042}, 
a C compiler used by many companies\footnote{https://www.absint.com/partners.htm}, 
Fiat-Crypto \cite{8835346} 
used in BoringSSL --a cryptographic library used in Chrome, Android, 
and CloudFlare--, ConCert \cite{10.1145/3372885.3373829}, 
a smart contract certification framework, etc. \textbf{FIXME: More examples}

@@

In Coq, \emph{Theorem} is the keyword used to define a specification, 
followed by the name (\emph{plus\_comm}). This specifies that for any 
two natural numbers $m$ and $n$ (encoded as \texttt{forall n m}), 
adding $n$ to $m$ is the same as adding $m$ to $n$ 
(encoded as \texttt{plus n m = plus m n}).
To write a formal proof of this statement, we start by indicating that 
we will use the tactic language with the keyword \emph{Proof}. Once the 
proof is complete, we use the keyword \emph{Qed} to signal to Coq that 
our proof is finished. At this point, Coq verifies that the proof is complete 
and correct according to its logic. If the proof is not finished, Coq will 
not accept \emph{Qed}. Therefore, having a \emph{Qed} for any specification 
ensures that our proof is correct, provided we trust the Coq theorem prover.
However, at this point, the curious reader can argue that why should we trust 
Coq because it is a computer program, similar any other computer program? 
Trust in Coq relies on two factors: (i) its meta-theory, the 
\emph{Calculus of Inductive Constructions} \cite{thierry1988calculus}, and 
(ii) its implementation in OCaml \cite{the_coq_development_team}.
The meta theory of Coq has been scrutinised for decades
by many mathematicians and so far it has withstood all 
the scrutiny, but not so much for its OCaml implementation \cite{coq_critical_bugs}. 
Therefore, there is a recent effort to develop more 
rigorous implementation for Coq proof checking \cite{10.1145/3371076} \textbf{FIXME: this is the place 
to state that we are following this path of MetaCoq and extraction}.

@@

In the Schnorr protocol, the elements $g$, $h$, and $a$ are group elements from the underlying group, while $u$, $c$, and $r$
are scalar values from the underlying prime field. In practice, both the group elements and the field elements are large 
numbers modulo some prime numbers, although the operations on them differ. 
One possibility is to encode the Coq functions to work with concrete numbers from the underlying group and field, 
but this leads to cumbersome proofs because of dealing with concrete numbers, but the bigger problem 
is that it is not modular. For most programming languages, this distinction 
may not be significant, but it can be a potential source of bugs \cite{10.1007/978-3-662-63958-0_24}. 
However, when encoding it in Coq, we need to thoroughly detail these distinctions because 
the same large number can belong to a group or field and that dictates what operations are available for them. 
Therefore, to avoid this unpleasant situation,  
we decided to abstract the underlying group as an abstract type $G$, endowed with 
usual group operation, and the underlying field as an abstract type $F$, endowed with 
filed operation in our formalisation. Moreover, we abstract these two types into a vector space for smooth interaction between 
group element and field element (exponential needs this). This 
makes our formalisation generic that can be instantiated with 
any kind of group and field as long as it follows the axioms of 
vector space. For example, in our formalisation we have developed 
efficient algorithm for the Schnorr group but our formalisation can 
later be also instantiated with elliptic curve group. However, 
one of the major benefit of this abstraction is easiness of proofs 
which can be easily automated. Therefore, we model the Schnorr 
protocol, the simulator, and the verification equation in Coq as follows:

@@

\noindent
 where $F$ is the underlying field, $G$ is the underlying group,
 $+ : F \rightarrow F \rightarrow F$ is the field addition,
 $* : F \rightarrow F \rightarrow F$ is the field multiplication, 
 $gop : G \rightarrow G \rightarrow G$ is the group operation, and
 $\mbox{\textasciicircum}: G \rightarrow F \rightarrow G$ is the scalar multiplication of vector space.  
 Recall that a vector space consists of a set of 
 vectors, denoted by \(V\), a field \(F\) (elements of F are called scalars), and 
 two operations: vector addition and scalar multiplication.
\begin{itemize}
    \item \textit{Vector addition}: for any vectors \(\mathbf{v}, \mathbf{w} \in V\), their sum, denoted by \(\mathbf{v} + \mathbf{w}\), is also in \(V\).
    \item \textit{Scalar multiplication}: for any vector \(\mathbf{v} \in V\) and any scalar \(c \in F\), their product, denoted by \(c\mathbf{v}\), is also in \(V\).
\end{itemize}

@@

We prove the completeness and special soundness property of the Schnorr protocol, 
as shown in Listing \cite{comp_sound}. However, to prove the special honest-verifier 
zero-knowledge property we need a model to reason about uniform probability 
distribution in Coq. We discuss the encoding of uniform probability distribution 
in the next section and subsequently explain the special honest-verifier zero-knowledge
proof.

@@

\subsection{Uniform Distribution}
To reason about special honest-verifier zero-knowledge proof 
intuitively, we need a model of uniform distribution. 
We follow the \cite{erwig2006functional}  and 
encode the distribution a list of tuples where the 
second element of tuple basically is probability of first element 
in the tuple. Moreover, we define equality on distribution by 
stating that two distributions are equal if they are permutation 
of each other. This encoding is flexible and allows us 
to encode any kind of distribution. We prove that the dist is a 
monad \cite{erwig2006functional} by 
giving it a suitable definitions of ret and bind. Making 
it an instance of monad makes our life easier because 
we can compose a distribution $d$ $n$ times to 
returns a distribution on a vector of $n$ elements. 
However, to reason about cryptographic primitives, we need uniform 
distribution and therefore we define a function, $uniform\_with\_replacement$, 
to generate uniform distributions, shown in Listing \ref{prob_def}.
We prove various prop

@@

Now we have all the ingredients to prove special honest-verifier 
zero-knowledge proof of the Schnorr protocol. We define the 
Schnorr distribution $schnorr\_distribution$
that involves the secret $x$ and 
the simulator distribution $simulator\_distribution$
that does not involve the secret $x$. 
In the Schnorr distibution we draw 
a random field element $u$ from the uniform 
distribution $uniform\_with\_replacement \text{ }lf \text{ } Hlfn$,
where $lf$ is list of elements from which $u$
is drawn and $Hlf$ is a proposition that 
$lf$ is non-empty, and use it to construct 
a distribution on sigma protocol. Similarly, 
for the simulator distribution 
we draw a random field element $u$ from 
$uniform\_with\_replacement \text{ }lf \text{ } Hlfn$
and use it construct a distribution on sigma protocol. 
Using these two definitions, we prove that these two 
distributions are identical 
$special\_honest\_verifier\_zkp$. However, 
contrary to the reasoning in cryptographic papers, 
these two distributions are not syntactically equal but 
semantically equal. For both distributions, 
the first component is a transcript $(a, c, r)$ and 
the second component is a rational number that represents 
its probability. The triples in both distrubtions 
are an accepting transcript (semantic) but they 
are not identical and differ syntactically; 
therefore we first apply the 
$accepting\_conversation$ to the first pair of 
both distributions which turn them into 
a boolean value (true). After this  
application both distributions are identical 
that we prove.

\begin{lstlisting}[frame=single, language=Coq, caption={Definition of Natural Number},
  label={ind_nat},captionpos=t, basicstyle=\ttfamily\footnotesize,
  abovecaptionskip=-\medskipamount]
(* Distribution that involves the secret x *)
Definition schnorr_distribution  (lf : list F) 
  (Hlfn : lf <> List.nil) (x : F) (g : G) (c : F) : 
  dist sigma_proto :=
  (* draw u from a random distribution *)
  u <- (uniform_with_replacement lf Hlfn) ;;
  Ret (schnorr_protocol x g u c).

@@

In the previous section, we explained the Schnorr proof, 
but we have also formalised a way to combine $n$ And, Or, Eq, 
and Parallel Schnorr statements. Proving special 
honest-verifier zero-knowledge proof is more challenging 
than a single Schnorr statement. One of the major challenge 
is to draw $n$ random element for an uniform distribution 
and therefore we have written the function $repeat\_dist\_ntimes\_vector$
that takes a distribution $dist\text{ }A$ and a number $n$ and a returns 
distribution of vectors of length $n$,  encoded as 
$dist \text{ } (Vector.t \text{ } A \text{ } n)$. The 
reader may notice that it is a generic function 
where distribution can be instantiated with 
any kind of distribution. However, in cryptography 
we are mostly concerned with uniform distribution and  
therefore in $uniform\_probability\_multidraw\_prob$ theorem
we instantiate the distribution with 
the uniform distribution $(uniform\_with\_replacement\text{ }lf\text{ }Hlf)$ and 
prove that the probability of this distribution 
is $1/|lf|^n$. We use it 
in Parallel composition where we combine $n$ sigma protocol and  
we need to need to draw a vector $n$ random numbers to 
prove the zero-knowledge property, as shown 
in $parallel\_schnorr\_distribution$.

\begin{lstlisting}[frame=single, language=Coq, caption={Definition of Natural Number},
label={ind_nat},captionpos=t, basicstyle=\ttfamily\footnotesize,
abovecaptionskip=-\medskipamount]
Fixpoint repeat_dist_ntimes_vector {A : Type} 
(d : dist A) (n : nat) : dist (Vector.t A n) := 
match n with 
| 0 => Ret (@Vector.nil A)
| S n' => 
  Bind d (fun u => 
    Bind (repeat_dist_ntimes_vector d n')
    (fun v => Ret (Vector.cons _ u _ v)))
end.

Lemma uniform_probability_multidraw_prob 
{A : Type} : forall n (lf : dist A) 
(a : Vector.t A n) b (Hlf : lf <> []), 
In (a, b) (repeat_dist_ntimes_vector 
  (uniform_with_replacement lf Hlf) n) ->
b = mk_prob 1 (Pos.of_nat 
(Nat.pow (List.length lf) n)).

@@

  (* Neutral Element *)
  Definition one : Schnorr_group.
    refine 
      (mk_schnorr 1 one_mod_p 
      one_pow_mod_q).
  Defined.

  Definition mul_schnorr_group 
      (u v : Schnorr_group) : Schnorr_group.
  (* Proof terms omitted *)
  Defined.

  (* u ^ (p - 2) is inverse of u *)
  Definition inv_schnorr_group 
  (u : Schnorr_group) : Schnorr_group.
  refine(
    match u with 
    | mk_schnorr au Hu Hv => mk_schnorr
        (Z.of_N (Npow_mod (Z.to_N au) 
        (Z.to_N (p - 2)) (Z.to_N p)))
        (@inv_schnorr_group_first au Hu)
        (inv_schnorr_group_second au Hu Hv)
    end).
  Defined.

>>>>

\usepackage{listings}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{listings}
\usepackage{tikz}
\usetikzlibrary{positioning, arrows.meta}
\usepackage{url}
\usepackage{xcolor}
\usepackage{soul}
\usepackage{lstcoq}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}
\title{Bridging Frontend and Backend: Implementing Sigma Protocol in WebAssembly and Rust from a Single Coq Formalisation}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract
\begin{abstract}
  \st{Despite the recent growth of zero-knowledge succinct argument 
  of knowledge (ZkSNARK)}, 
  $\Sigma$-protocol remains a popular choice in electronic voting to maintain privacy 
  and verifiability of elections and still used in prominent electronic voting systems such as 
  Helios, Belenois, SwissPost, etc. However, 
  the existing implementations of sigma protocol have two 
  shortcomings: (i) the same code is written twice, once for the front-end (JavaScript/TypeScript)
  and once for the back-end (Python/OCaml/Java), which is time-consuming but more importantly error-prone, 
  and (ii) none of these implementations offer any security guarantees.

@@

\subsection{Outline}
The rest of the paper is organised as follows: in section \ref{sigma_protocol}
we explain the idea of zero-knowledge proof and sigma protocol. 
Section \ref{coq_theorem} contains a brief introduction to 
the Coq theorem prover, MetaCoq, Coq encoding of the Schnorr protocol, 
uniform probability distribution model, modular arithmetic, Fiat-Shamir transform,
and reasoning about the complexity of extractor and simulator. 
In section \ref{case_studies} we show how our library can be used 
for encoding $\Sigma$-protocol used in Helios, Belenois, and SwissPost. 
Finally, section \ref{rel_work} concludes the related work, limitation of 
our work, and future direction.

\section{Sigma Protocol}\label{sigma_protocol}
Despite the recent growth and popularity of zero-knowledge succinct 
argument of knowledge (ZkSNARK), $\Sigma$-protocol remains a
leading cryptographic proof system in privacy-preserving voting systems. 
$\Sigma$-protocol is used in Helios --a widely popular voting system-- \cite{adida2008helios}, 
Belenois --used in French elections for overseas 
voters\footnote{VVFE, used in French elections, is a derivative of (part of) Belenios and 
inherited all the zero-knowledge proofs from Belenois.}-- \cite{cortier2023french}, and 
SwissPost --used in Swiss elections for overseas voters-- voting 
systems \cite{10.1007/978-3-031-15911-4_4} (\textcolor{red}{Look for more examples}). 
Moreover, recently zero-knowledge proof community started a movement to standardise
the sigma protocol \cite{ZKProof}.
One of the main reasons for its popularity in privacy-preserving voting systems is its
efficiency and well-understood security.  Other application of sigma protocol 
include anonymous credentials \cite{10.1145/2660267.2660328}, 
password-authenticated key exchange \cite{10.5555/2022815.2022838}, 
signature \cite{10.1007/0-387-34805-0_22}, threshold signature \cite{10.1007/978-3-030-81652-0_2}, 
etc.

%\textbf{Introduce here that sigma protocol is efficient and simple and all other details about sigma protocol.} 
Zero-knowledge proofs were first studied by Goldwasser, Micali, and Rackoff \cite{10.1145/22145.22178} and 
are possible for all problems in $NP$ \cite{10.1145/116825.116852}. Recall that $NP$ is the class of 
problems that have efficient verifiers; that is, for a given problem in the $NP$ class, there is 
a polynomial-time algorithm that can verify whether a given solution, or witness, to the problem is correct or not.
More formally, for a given binary $NP$-relation $R$ and a statement $x$, zero-knowledge proof allows a prover 
$P$ to convince a verifier $V$ that they posses a witness $w$ --polynomial in length of $x$-- such 
that ($x$, $w$) $\in$ $R$. To achieve this, $P$ and $V$ engage in a 
interactive protocol and  at the end of the protocol $V$ either accepts or rejects the proofs.
$\Sigma$-protocol --shown in the Figure \ref{fig:sig_picture}-- is an efficient three move zero-knowledge proof where the 
verifier is assumed to be honest and characterised by completeness, (special) soundness, and 
(special honest-verifier) zero-knowledge. More formally: 
\begin{itemize}
  \item \textbf{Completeness:}
  when $P$ and $V$ follow the protocol, $V$ always accepts. 
  In other words, if $P$ knows $w$ for $x$ such that ($x$, $w$) $\in$ $R$,  
  then $P$ can successfully construct a proof $(a, c, r)$ that always passes the verification equation, 
  i.e., $V$ accepts.

\item \textbf{Special soundness:}
there exists a polynomial-time extractor $E$ which given 
two accepting transcripts $(a, c, r)$ and $(a, c', r')$ for $x$, 
it can extract a witness $w$ such that $(x, w) \in R$. In other words, 
if $P$ does not know $w$ for $x$ then it cannot construct a proof 
$(a, c, r)$ that will passes the verification check. 
\item \textbf{Special honest-verifier zero-knowledge proof:}
There exists a polynomial-time simulator $S$ that, for any 
$v \in L_R$ where $L_R = \{x  \mid (x, w) \in R\}$,
generates an accepting transcript $(a, c, r)$ 
without using the witness $w$. 
This transcript is indistinguishable from one produced by an 
honest conversation between $P$ and $V$ using $w$, 
satisfying $(v,w) \in R$.

@@

Sigma protocol was first defined by Ronald Cramer \cite{cramer1996modular} 
and the first efficient sigma protocol was introduced by Schnorr in \cite{schnorr1991efficient}. 
Even though there exists many sigma protocols, the Schnorr protocol is used extensively in privacy-preserving voting systems, e.g, Helios, Belenois, SwissPost, etc. Therefore, 
our formalisation is geared towards the Schnorr protocol. 
Moreover, we also formalise the composition of (i) $n$ Parallel Schnorr, 
(i) $n$ And Schnorr, (iii) $n$ Or Schnorr, and (iv) $n$ Eq Schnorr protocol, for an arbitrary natural number $n$. 
Ideally, given the abstraction of our formlisation (\textit{vector space}), we could have 
abstracted the sigma protocol and their composition in a generic way \cite{10.1007/978-3-642-02384-2_17},
but we decided to focus on the Schnorr protocol and its composition to ensure familiar APIs 
for electronic voting in WebAssembly and Rust. We also formalise Okamoto protocol as a sub-protocol 
for Neq composition (\textcolor{red}{Berry: does it sound okay?}). However, we have not seen 
much use of Neq in the privacy-preserving voting systems, at least in the existing voting systems 
such as Helios, Belenois, and SwissPost.

\section{The Coq Theorem Prover}\label{coq_theorem}
\textcolor{red}{FIXME:this seems too much of tutorial Coq so let me know your thoughts. We 
can explain MetaCoq here as well and the extraction to Rust and WebAssembly
so that it justifies this tutorial}
The Coq theorem prover \cite{the_coq_development_team} is a computer program that interacts with users, 
enabling them to encode mathematical definitions, express specifications (true statements) about 
their definitions, and formally prove that the definitions imply the specifications. 
A user first defines their mathematical object (definition) and then figures 
out some true statements (specifications) about 
that object. Once the user has definition and specifications, they need to prove 
that the definition satisfies the specifications 
(proof writing). In general, amongst these three steps, 
the proof writing step is the most challenging and important step. 
During the proof writing, an user interacts with the Coq theorem prover 
via \emph{tactics}, a domain specific language 
to ease the proof writing. Even though the Coq theorem 
provides some amount of automation and proof 
search, most of the time it requires human assistance to finish the proof. 
It is one the most popular and robust theorem, and
one of reasons for its popularity is 
dependent types, that makes it possible to encode
any mathematical statement in its logic.
It has been used to verify many real world software projects and mathematical 
artefacts. For example, Coq has been used to 
formally verify CompCert \cite{10.1145/1111037.1111042}, 
a C compiler used by many companies\footnote{https://www.absint.com/partners.htm}, 
Fiat-Crypto \cite{8835346} 
used in BoringSSL --a cryptographic library used in Chrome, Android, 
and CloudFlare--, ConCert \cite{10.1145/3372885.3373829}, 
a smart contract certification framework, etc. \textcolor{red}{FIXME: More examples}

@@

In Coq, \emph{Theorem} is the keyword used to define a specification, 
followed by the name (\emph{plus\_comm}). This specifies that for any 
two natural numbers $m$ and $n$ (encoded as \texttt{forall n m}), 
adding $n$ to $m$ is the same as adding $m$ to $n$ 
(encoded as \texttt{plus n m = plus m n}).
To write a formal proof of this statement, we start by indicating that 
we will use the tactic language with the keyword \emph{Proof}. Once the 
proof is complete, we use the keyword \emph{Qed} to signal to Coq that 
our proof is finished. At this point, Coq verifies that the proof is complete 
and correct according to its logic. If the proof is not finished, Coq will 
not accept \emph{Qed}. Therefore, having a \emph{Qed} for any specification 
ensures that our proof is correct, provided we trust the Coq theorem prover.
However, at this point, the curious reader can argue that why should we trust 
Coq because it is a computer program, similar any other computer program? 
Trust in Coq relies on two factors: (i) its meta-theory, the 
\emph{Calculus of Inductive Constructions} \cite{thierry1988calculus}, and 
(ii) its implementation in OCaml \cite{the_coq_development_team}.
The meta theory of Coq has been scrutinised for decades
by many mathematicians and so far it has withstood all 
the scrutiny, but not so much for its OCaml implementation \cite{coq_critical_bugs}. 
Therefore, there is a recent effort to develop more 
rigorous implementation for Coq proof checking \cite{10.1145/3371076} \textcolor{red}{FIXME: this is the place 
to state that we are following this path of MetaCoq and extraction}.

@@

In the Schnorr protocol, the elements $g$, $h$, and $a$ are group elements from the underlying group, while $u$, $c$, and $r$
are scalar values from the underlying prime field. In practice, both the group elements and the field elements are large 
numbers modulo some prime numbers, although the operations on them differ. 
One possibility is to encode the Coq functions to work with concrete numbers from the underlying group and field, 
but this leads to cumbersome proofs because we need to deal with the details of concrete numbers during the security proofs. 
More importantly, it is not modular because it binds the formalisation to multiplicative groups; therefore, 
it can no longer be used over elliptic curve groups. For most programming languages, this distinction 
may not be significant, but it can be a potential source of bugs \cite{10.1007/978-3-662-63958-0_24} undermining 
the security of underlying cryptographic primitives. 
However, when encoding it in Coq, we need to thoroughly detail these distinctions because 
the same large number can belong to a group or field and that dictates what operations are available for them. 
Therefore, to avoid this unpleasant situation,  
we abstract the underlying group as an abstract type $G$, endowed with 
usual group operation, and the underlying field as an abstract type $F$, endowed with 
field operation in our formalisation. 
Moreover, we abstract these two types into a vector space for smooth interaction between 
group element and field element (exponentiation in multiplicative group or mulitplication of a field element 
with a curve point in ellptic curve group is an instance of this interaction). This 
makes our formalisation generic that can be instantiated with 
any kind of group and field as long as it follows the axioms of 
vector space. For example, in our formalisation we have developed 
efficient algorithm for the Schnorr (mulitplicative) group but our formalisation can 
later be also instantiated with elliptic curve group. However, 
one of the major benefit of this abstraction is easiness of proofs 
which can be easily automated. Therefore, we model the Schnorr 
protocol, the simulator, and the verification equation in Coq as follows:

@@

\noindent
 $F$ is the underlying field, $G$ is the underlying group,
 $+ : F \rightarrow F \rightarrow F$ is the field addition,
 $* : F \rightarrow F \rightarrow F$ is the field multiplication, 
 $gop : G \rightarrow G \rightarrow G$ is the group operation, and
 $\mbox{\textasciicircum}: G \rightarrow F \rightarrow G$ is the scalar multiplication of vector space.  
 Recall that a vector space consists of a set of 
 vectors, denoted by \(V\), a field \(F\) (elements of F are called scalars), and 
 two operations: vector addition and scalar multiplication.
\begin{itemize}
    \item \textit{Vector addition}: for any vectors \(\mathbf{v}, \mathbf{w} \in V\), their sum, denoted by \(\mathbf{v} + \mathbf{w}\), is also in \(V\).
    \item \textit{Scalar multiplication}: for any vector \(\mathbf{v} \in V\) and any scalar \(c \in F\), their product, denoted by \(c\mathbf{v}\), is also in \(V\).
\end{itemize}

@@

We prove the completeness and special soundness property of the Schnorr protocol, 
as shown in Listing \ref{comp_sound}. However, to prove the special honest-verifier 
zero-knowledge property we need a model to reason about uniform probability 
distribution in Coq. We discuss the encoding of uniform probability distribution 
in the next section and subsequently explain the special honest-verifier zero-knowledge
proof.

@@

\subsection{Uniform Distribution}
To reason about special honest-verifier zero-knowledge proof 
intuitively, we need a model of uniform distribution. 
We follow \cite{erwig2006functional}  and 
encode the distribution a list of tuples where the 
second element of a tuple is probability --represented as 
a rational number-- of the first element 
of the tuple. Moreover, we define equality on distribution by 
stating that two distributions are equal if they are permutation 
of each other. This encoding is flexible and allows us 
to encode any kind of distribution. We prove that the dist is a 
monad \cite{erwig2006functional} by 
giving it a suitable definitions of \textit{ret} and 
\textit{bind}. Making it an instance of monad makes our life easier because 
we can compose a distribution $d$ $n$ times to 
returns a distribution on a vector of $n$ elements, 
which is needed in proving the special honest-verifier 
zero-knowledge proof of $n$ composed $\Sigma$-protocol.
However, we need uniform 
distribution, rather than an arbitrary distribution, to reason about 
zero-knowledge of $\Sigma$-protocol, and therefore we define a function, $uniform\_with\_replacement$, 
to generate uniform distributions, shown in Listing \ref{prob_def}.

@@

Now we have all the ingredients to prove special honest-verifier 
zero-knowledge proof of the Schnorr protocol. We define the 
Schnorr distribution $schnorr\_distribution$
that involves the secret $x$ and 
the simulator distribution $simulator\_distribution$
that does not involve the secret $x$, shown in Listing \ref{def_zero}. 
In the Schnorr distibution we draw 
a random field element $u$ from the uniform 
distribution $uniform\_with\_replacement \text{ }lf \text{ } Hlfn$,
where $lf$ is list of elements from which $u$
is drawn and $Hlf$ is a proposition that 
$lf$ is non-empty, and use it to construct 
a distribution on sigma protocol. Similarly, 
for the simulator distribution 
we draw a random field element $u$ from 
$uniform\_with\_replacement \text{ }lf \text{ } Hlfn$
and use it construct a distribution on sigma protocol. 
Using these two definitions, we prove that these two 
distributions are identical $special\_honest\_verifier\_zkp$. However, 
contrary to the reasoning in cryptographic papers, 
these two distributions are not syntactically equal but 
semantically equal. For both distributions, 
the first component is a transcript $(a, c, r)$ and 
the second component is a rational number, which represents probability of 
the transcript $(a, c, r)$. The triples in both distrubtions 
are an accepting transcript (semantic) but they 
are not identical and differ syntactically; 
therefore we first apply the 
$accepting\_conversation$ to the first pair of 
both distributions which turn them into 
a boolean value (true). After this  
application both distributions are identical 
that we prove.

\begin{lstlisting}[frame=single, language=Coq, caption={Special Honest-Verifier Zero-Knowledge Proof},
  label={def_zero},captionpos=t, basicstyle=\ttfamily\footnotesize,
  abovecaptionskip=-\medskipamount]
(* Distribution that involves the secret x *)
Definition schnorr_distribution  (lf : list F) 
  (Hlfn : lf <> List.nil) (x : F) (g : G) (c : F) : 
  dist sigma_proto :=
  (* draw u from a random distribution *)
  u <- (uniform_with_replacement lf Hlfn) ;;
  Ret (schnorr_protocol x g u c).

@@

In the previous section, we explained the Schnorr proof, 
but we have also formalised a way to combine $n$ And, Or, Eq, 
and Parallel Schnorr statements and proving their special 
honest-verifier zero-knowledge proof is more challenging 
than a single Schnorr statement. One of the major challenge 
is to draw $n$ random element for an uniform distribution 
and therefore we have written the function $repeat\_dist\_ntimes\_vector$ --shown in Listing \ref{dist_comp}--
that takes a distribution $d$ and a number $n$ and a returns 
distribution of vectors of length $n$,  encoded as 
$dist \text{ } (Vector.t \text{ } A \text{ } n)$. The 
reader may notice that it is a generic function 
where distribution can be instantiated with 
any kind of distribution. However, in cryptography 
we are mostly concerned with uniform distribution and  
therefore in $uniform\_probability\_multidraw\_prob$ theorem
we instantiate the distribution with 
the uniform distribution $(uniform\_with\_replacement\text{ }lf\text{ }Hlf)$ and 
prove that the probability of this distribution 
is $1/|lf|^n$. We use it in Parallel composition where we combine $n$ statments and  
we need to draw a vector of $n$ random field elements to 
prove the zero-knowledge property, as shown in $parallel\_schnorr\_distribution$.

\begin{lstlisting}[frame=single, language=Coq, caption={Composition of a Distribution},
label={dist_comp},captionpos=t, basicstyle=\ttfamily\footnotesize,
abovecaptionskip=-\medskipamount]
Fixpoint repeat_dist_ntimes_vector {A : Type} 
(d : dist A) (n : nat) : dist (Vector.t A n) := 
match n with 
| 0 => Ret (@Vector.nil A)
| S n' => 
  Bind d (fun u => 
    Bind (repeat_dist_ntimes_vector d n')
    (fun v => Ret (Vector.cons _ u _ v)))
end.

Lemma uniform_probability_multidraw_prob 
{A : Type} : forall n (lf : dist A) 
(a : Vector.t A n) b (Hlf : lf <> []), 
In (a, b) (repeat_dist_ntimes_vector 
  (uniform_with_replacement lf Hlf) n) ->
b = mk_prob 1 (Nat.pow (List.length lf) n).

@@

  (* u ^ (p - 2) is inverse of u *)
  Definition inv_schnorr_group 
  (u : Schnorr_group) : Schnorr_group.
  refine(
    match u with 
    | mk_schnorr au Hu Hv => mk_schnorr
        (* terms omitted *)
    end).
  Defined.
